---
title: "Practical Machine Learning"
author: "Koh Tongyi"
date: "19 August 2015"
output: html_document
---

###Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset). 

###Data 
The training data for this project are available here: 

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>

The test data are available here: 

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>

The data for this project come from this source: <http://groupware.les.inf.puc-rio.br/har>.

###Submission

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

1. Your submission should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).
2. You should also apply your machine learning algorithm to the 20 test cases available in the test data above. Please submit your predictions in appropriate format to the programming assignment for automated grading. See the programming assignment for additional details. 

###Reproducibility 

Load required libraries and set seed
```{r}

library(caret)
library(randomForest)
library(rpart)

set.seed(1235) # set seed
```

### Loading and Cleaning data
Download the training and test csv data and place them in the same directory of the script.
```{r}
naValues<-c("", "NA", "#DIV/0!")
trainingSet <- read.csv("pml-training.csv", na.strings = naValues)
testingSet <- read.csv("pml-testing.csv", na.strings = naValues)

#Remove var not suitable for prediction
trainingSet <- trainingSet[, -(1:7)]

#Using nearZeroVar function can be used to identify near zero-variance variables 
nearZero<-nearZeroVar(trainingSet, saveMetrics = TRUE)
trainingSet <- trainingSet[, !nearZero$nzv]

#Remove columns if number of row with NA is >50% 
trainingSet <-trainingSet[, colSums(is.na(trainingSet)) < (nrow(trainingSet)*0.5)]

#Create training set and test set for cross validation
partition <- createDataPartition(y=trainingSet$classe, p=0.7, list=FALSE)
myTrainingSet <- trainingSet[partition, ] 
myTestingSet <- trainingSet[-partition, ]

```

### Model selection
We compare 2 model: Decision Tree and Random Forest.

####Decision Tree
```{r}
dt <- rpart(classe ~ ., data=myTrainingSet, method="class")

dtPrediction <- predict(dt, myTestingSet, type = "class")

confusionMatrix(dtPrediction, myTestingSet$classe)
```

####Random Forest
```{r}
rf <- randomForest(classe ~ ., data=myTrainingSet, method="class")

rfPrediction <- predict(rf, myTestingSet, type = "class")

confusionMatrix(rfPrediction, myTestingSet$classe)
```

####Conclusion
The training set is partitioned to perform cross-validation. The result shows that Random forest model(0.9951) has a higher accuracy than Decision Tree(0.7468). The out-of-sample error is estimated at 0.005, or 0.5%, calculated with 1 - accuracy for predictions. With such high accuracy, Random Forest is choosen for the submission.

###Submission
```{r}
result <- predict(rf, testingSet, type="class")

result

#Write files
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(result)
```
